---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abetareg

<!-- badges: start -->
<!-- badges: end -->

## Loglikelihood Adjustment for Beta Regression Models

The *abetareg* package adjusts the log-likelihood function for beta regression from the [betareg](https://cran.r-project.org/package=betareg) package to provide robust sandwich estimation of parameter covariance matrix. The methodology of the adjustment is based on the [chandwich](https://cran.r-project.org/package=chandwich) package.

## An Example

```{r}
library(abetareg)
data("sp", package = "abetareg")
# Data scaling suggested by Smithson and Verkuilen (2006)
x <- sp$performance
x1 <- (x - 10) / (100 - 10)
x2 <- (x1 * (length(x) - 1) + 0.5) / length(x)
sp$performance <- x2
# Fit a beta regression model for student performance
library(betareg)
opt_naive <- betareg(performance ~ studyh + ea | sleeph, data = sp)
# Adjust the log-likelihood function
adj_naive <- alogLik(opt_naive)
(s1 <- summary(adj_naive))
# Account for the clustered structure
adj_naive2 <- alogLik(opt_naive, cluster = sp$sqpp)
(s2 <- summary(adj_naive2))
```

### Installation

To get the current released version from CRAN:

``` {r, warning = FALSE}
install.packages("abetareg")
```

### Vignette

See `vignette("abetareg", package = "abetareg")` for an overview of the package.
